Beam test performance of the highly granular SiW-ECAL technological prototype for the ILC.

### 2018-06-12 performance-paper-draft/shortversion/draft.tex
First draft of the short version

\usepackage[utf8]{inputenc}
\usepackage{textcomp}
\usepackage{lineno}
\usepackage{color}
\newcommand{\todo}[1]{\textcolor{red}{{#1}}}

@@

\begin{frontmatter}
  \title{Commissioning and performance in beam tests of the highly granular SiW-ECAL technological prototype for the ILC.}

@@

  \end{frontmatter}
\linenumbers

@@

\subsection{Silicon sensors}
\label{sec:wafers}

The sensors consist on floating zone silicon wafers 320$\mu$m thick with high resistivity (bigger than 5000 $\Omega\cdot$cm).
The size of the wafer is $9\times9$ cm$^{2}$ and it is subdivided in an array of 256 PIN diodes of $5\times5$ mm$^{2}$.
A MIP traversing the PIN parallel to its normal will create $\sim$ 80 $h^{+}e^{-}$ pairs per $\mu$m which corresponds to 4.1 fC
for particles incident perpendicularly to its surface.

The original design of the silicon wafers included an edge termination made of floating guard-rings.
It was observed in beam tests \cite{Cornat:2015eoa,Cornat:2009zz} that the capacitive coupling between such floating guard-rings 
and the channels at the edge was not negligible in tests with high energy beams (pions and electrons with energies larger than 20-40 GeV).
This coupling lead to fake events in which, at least,
the channels in the four edges of the wafer are triggered at the same time. 
This is why these events are called squared events. 
An R\&D program together with Hamamatsu Photonics (HPK Japan) was conducted to study the guard-rings design 
as well as the internal crosstalk.
%As outcome of this R\&D program, the technological prototype 
%is equipped with silicon wafers with:
It was concluded that using wafers without guard rings and with a width of the peripheral areas lower than 
500 $\mu$m thanks to the use of stealth dicing technique, the amount of these squared events 
can be reduced to be almost negligible. This need to be confirmed in beam test.
Unfortunately, for the interaction with low energy particles as the delivered at the
DESY beam test facility (see Section \ref{sec:beamtest}) the
amount of squared events is expected to be negligible, therefore we will not discuss this issue in the following.
% do any differentiation 
%of our results depending on the wafer type mounted in the detector.

%\begin{itemize}
%\item \todo{N wafers type-X with guard-rings of XXX characteristics (continuous, segmented, segmented smaller??)}
%\item \todo{N wafers type-Y without guard rings and the width of the peripheral dead areas lower than 500 $\mu$m thanks to the use of stealth dicing technique.}
%\end{itemize}

%For the wafers \todo{type-Y}, the amount of squared events is expected to be reduced by, at least, a \todo{factor XXX}. In both cases, 
%for interaction with low energy particles as the delivered at the DESY beam test facility (see Section \ref{sec:beamtest}) the
%amount of squared events is expected to be negligible, therefore we will not do any differentiation 
%of our results depending on the wafer type mounted in the detector.

%\todo{ Is this needed? If so, what are the right plots to show?} Before the assembly of the sensors in the detector, these are characterized one-by-one by measuring the leakage current (of %the order of 1 nA per single diode) 
%as a function of the bias voltage (I-V curve), the full depletion voltage (of 40 V) extracted from C-V curve and
%the stability in time (leakage current versus time at a nominal bias).
%The single diodes have capacitances of 8 $pF$.

%\begin{figure}[!t]
%\centering
%\begin{tabular}{lll}
%  \includegraphics[width=1.8in]{../figs/test.jpg} &
%  \includegraphics[width=1.8in]{../figs/test2.jpg} &
%  \includegraphics[width=1.8in]{../figs/test3.jpg}
%\end{tabular}
%\caption{I-V curve, C-V curve, leakage current vs time. \todo{Temporary picture: Triptic La Esperanza del Condenado, J. Miró}}
%\label{shortslab}
%\end{figure}

\subsection{SKIROC: Silicon pin Kalorimeter Integrated ReadOut Chip}
\label{sec:skiroc}

The SKIROC\cite{Callier:2011zz} (Silicon pin Kalorimeter Integrated ReadOut Chip) is a
very front end ASIC (application-specific integrated circuits)
designed for the readout of the Silicon PIN didoes.
In its version SKIROC2 it consists of 64 channels in AMS 0.35 $\mu$m SiGe technology.
Each channel comprises a low noise charge preamplifier of variable gain followed by two branches:
a fast shaper for the trigger decision and a set of dual gain slow shaper for charge measurement. 
The gain can be controlled by modifying the feedback capacitance during the configuration of the detector.
With the slowest gain, 6pF, the ASIC will handle a linear dynamic range from 0.1 to up to 1500 MIPs 
(a slightly less than the desired final value for the ILC). 
Finally, a Wilkinson type analogue to digital converter fabricates the digitized charge deposition that can be readout. 
Once one channel is triggered, the ASIC reads out all 64 channels adding a bit of information to tag them as
triggered or not triggered and the information is stored in 15 cell deep physical switched capacitor array (SCA).

The SKIROC ASICs can be power-pulsed by taking advantage of the ILC spill structure: 
the bias currents of the ASIC can be switched off during the idle time between bunch trains.
With this method, the ASIC is able to reduce its power consumption down to 25 $\mu$W per channel,
meeting the ILC requirements. All the results shown in this paper are obtained in power pulsing mode.

\begin{figure}[!t]
  \centering
    \includegraphics[width=4in]{../figs/skiroc2_block.eps}
\caption{The schematics of the analog part of SKIROC2. \todo{High-stack picture (right bottom corner)}}
\label{SKIROC2}
\end{figure}

\subsection{Active Sensor Units}
\label{sec:ASU}

\begin{figure}[!t]
  \centering
  \begin{tabular}{l}
    \includegraphics[width=4in]{../figs/ASU_geometry1.eps}  \\
    \includegraphics[width=4in]{../figs/ASU_geometry2.eps}  \\
  \end{tabular}
  \caption{Repartition of the ASIC (up) and channels (down) in one ASU. In this perspective, the Si-Sensors are in glued in the back.
    The channels are separated (in x and y) by 5.5 mm.
    The empty cross in the middle of the ASU corresponds to the 1 mm separation between the sensors.
    The areas covered by the different ASICs and channels
    are labeled with numbers following design and DAQ criteria: from 0-16 in the case of the ASICs and from 0-63 in the case of the channels.
  }
\label{ASU}
\end{figure}

The entity
of sensors, thin PCB (printed circuit boards) and ASICs is called Active Signal Units or ASU.
An individual ASU has a lateral dimension of 18x18 cm$^{2}$.
The ASUs are currently equipped
further with 16 SKIROC2 ASICs for the read out and features 1024 square pads (64 per ASIC) of 5x5 mm.
The channels and ASICs are distributed along the ASU as shown in Figure \ref{ASU}. Each ASU is equipped with 4 silicon wafers as the described in Section \ref{sec:wafers}.
The high voltage is delivered to the wafers using a HV-kapton sheet that covers the full extension of the wafers.

\subsection{Data AcQuisition system}
\label{sec:DAQ}

The subsequent chain of the data acquisition (DAQ)\cite{Gastaldi:2014vaa} system is inspired by the ILC.
It consists on three modules.
The first module is the so called detector interface (DIF) which is placed at the beginning of each layer holding up to 15 ASUs;
All DIFs are connected by single HDMI cables to the concentrator cards that make
the second module: the Gigabit Concentrator Cards (GDCCs).
%The HDMI connection is used to transmit both slow control and data readout.
This cards are used to control up to 7 DIFs collecting all data from them and distributing among them the system clock and fast commands.
The last module, the most downstream, is the clock and control card (CCC) which
provides a clock, control fan-out of up to 8 GDCCs and accepts and distributes external signals (i.e. signals
generated external pulse generator to simulate the ILC spill conditions).
The whole system is controlled by the Calicoes and the Pyrame DAQ software version 3~\cite{Rubio-Roy:2017ere,Magniette:2018wdz}.
%The Pyrame framework provides basic blocks (called modules) of control-command or data acquisition.
%Calicoes is specific the implementation of these blocks for control-command and data acquisition of the SiW-ECAL prototype. 
%This new version of the DAQ software implements a new mechanism to allow any module to treat and publish data in real time. 
%Those data are made available to any requesting module, in particular to a new online monitor module developed
%for this beam test, allowing for quick online checks of the data quality and integrity.

\subsection{Readout layers and SLABs}
\label{sec:setup}

\begin{figure}[!t]
  \centering
    \includegraphics[width=2.8in]{../figs/test.jpg} 
  \caption{\todo{Temporary picture: La Esperanza del Condenado, J. Miró}: Open single SLAB with FEV11 ASU, 16 SKIROC 2, interface card and DIF visibles. }
\label{ASU2}
\end{figure}

The readout layers of the SiW-ECAL consist of a chain of ASUs and an adapter board
to a data acquisition system (DAQ) at the beginning of the layer.
This adapter board is called SMBv4
and it also serves as to hold other services as power connectors or the super capacitances used for the power pulsing. 
These capacitances of 400mF with 16 m$\Omega$ of equivalent serial resistance. 
The purpose of these capacitances is to provide local storage 
of the necessary charge to avoid the transport of current pulses over long cables, 
ensuring in this way the stability of the ASICs during the acquisition.

The readout layers are embedded on a "U" shape carbon structure to protect the wafers.
The full system is then covered by two aluminum plates
to provide electromagnetic shielding and mechanical stability.
This ensemble is denominated SLAB
(``short'' for 1 ASU ensembles or ``long'' for several ASUs enchained) and it can be seen in
Figure \ref{ASU2}.
With current SLABs, a potential density of
4000 channels/dm$^{3}$ is achievable. This number should be compared with
the density achieved for previous beam tests: 1500 channels/dm$^{3}$ \cite{Amjad:2014tha}.

\begin{figure}[!t]
\centering
\begin{tabular}{l}
\includegraphics[width=6.0in]{../figs/assembly.png} 
\end{tabular}
\caption{Process flow for the assembly of the SiW-ECAL SLABs.}
\label{assembly}
\end{figure}

The SiW-ECAL detector designed for the ILD requires of the order
of $10^{5}$ highly integrated detection like the ones described in this text.
For the production of the small sample of SLABs studied in this document,
a scalable working procedure has been established among several groups \cite{Boudry:2318814}
profiting from the funding of projects like AIDA2020 or the HIGHTEC emblematic project
of the P2IO. A schematic view of this assembly procedure chain can be seen in
Figure \ref{assembly}. For more details we refer to Ref.\cite{Boudry:2318814}.

\section{Commissioning}
\label{sec:commissioning}

This beam test was prepared by a careful and comprehensive commissioning comprising
the debug of the
short SLABSs with special emphasis in the control of the noise and the study of the
prototype performance in cosmic rays tests.

Earlier experiences with the SKIROC2 ASIC are reported in Refs. \cite{Amjad:2014tha,Suehara:2018mqk}). 
Internal SKIROC2 parameters found in these references are adopted in the following
except if the opposite is stated.
For example, the gain value of 1.2pF for the preamplifier is used. 
With this gain, the SKIROC2 ensures a linearity better than 90\% 
for 0.5-200 MIPs, which is enough for 
electromagnetic showers created by few GeV 
electrons or positrons.
% The delay of the trigger signal is the same than in 
%reference \cite{Amjad:2014tha,Suehara:2018mqk}: 130 DAC for all ASICS. 
%This value was chosen by selecting the delay in which an injected signal
%of few MIPs of value get the maximum value of ADC. %Finally, the value of the
%compensation capacitance at the output of the preamplifier has been set to 6pF instead of 
%4pF to prevent overshots and ripples of the signal used as input to the trigger
%decision and ADC lines.

\subsection{Optimization of noise levels}
\label{sec:comm_noise}

%The purpose of this phase was to monitor and control the
%noise levels of the prototype.
%The commissioning includes the definition of trigger threshold values and a list of noisy channels to be masked.
Studying and control the noise levels
is crucial since noisy channels may saturate the DAQ faster than physical signals.
Two different types of noise sources were identified:
a set of noisy channels randomly distributed in time 
and noise bursts affecting to all layers at the same time.
The outcome of this commissioning is summarized in Figure \ref{noisycells}.

The first type of noise source, noise events randomly distributed in time,
forced us to define a list of channels to be masked in all 
layers. Again, two different type of noisy channels have been identified. 
The first group is compose by randomly distributed in space channels.
The second group is made of
a fixed list of channels, all in the same positions for all SLABs,
and all associated to a higher rate of events with underflowed value
of the read ADC.
%The latter set of channels were identified by means of the online monitoring
%tools.
%As channels are located in the same place for all SLABs,
%we took the conservative assumption
%that all modules would have problems in the same set of channels.
%For example, all channels number 37.
A large amount of these channels are located
in areas of the PCB where the density of lines 
(data and power transmission) is higher than in others.
This hints for an issue on the routing of the PCB.
Therefore, these channels are tagged in Figure \ref{noisycells} with the "routing issues" label
although more tests and deeper inspections of the PCB layout are needed to clarify this issue.

The list of the noisy channels randomly distributed in time and space was
defined by means of dedicated data taking runs. 
These runs were characterized by: their short acquisition windows (open to data recording for 
only 1.1 ms) at low repetition frequencies (5 Hz) to minimize the chances of having real events due to cosmic rays 
hitting the detector during the
data taking; and the relatively high trigger threshold values between 250 and 400 DAC 
(which are equivalent to $\sim$0.5-2 MIP, see Section \ref{sec:comm_trigger} for more 
information). The full process was an iterative process starting with several repetitions
of runs at a high threshold value and then some iterations at lower threshold but always 
higher than 0.5 MIP. 
In each of the iterations, the channels identified as noisy were masked.
A full run involved 3-5 threshold values and at least two repetitions per value.
Again, we decided to follow a conservative approach for list the noisy channels to be masked:
if the channel was triggered at rates larger than 0.5-1\% of the total number of triggers per ASIC
it was added to the list. These channels are labeled as "other" in the first plot in  Figure \ref{noisycells}.

In addition to the different noisy channel types described above, we also have
masked full sectors of the SLABs if an ASIC was faulty (at least 70\% of channels 
listed as noisy) or if a Si-wafer was misworking (high leakage currents). In these cases 
and in all mentioned above, the masking involved two steps: disabling of the trigger and 
the disabling of the power of the preamplifier of that channel.

\begin{figure}[!t]
  \centering
  \begin{tabular}{l}
  \includegraphics[width=4in]{../figs/commissioning/masked_layer.eps} \\
  \includegraphics[width=4in]{../figs/commissioning/masked_chip.eps}
  \end{tabular}
\caption{Ratio of channels that are marked as noisy in all slabs. 
Top: inventory of the different type of noisy channels per slab. 
Bottom: break down of the total number of noisy channels per ASIC. 
The ASICs 4-7 (wafer issue) and 10 from layer 1 and the ASIC 4 from layer 7 are not included
in the second plot since they are fully masked.}
\label{noisycells}
\end{figure}

The other source of noise mentioned at the beginining of this section
consist on noise bursts happening coherently at the same time in all SLABS
and it was correlated to occasions when the electrical isolation between SLABs was broken.
This is a hint of a system effect as can be the appearance of
grounding loops or disturbances in the power supplies.
The issue was circumvented by improving the electrical isolation of single layers.
We have also observed that the noise bursts happens only at the end of long acquisitions. Therefore, in addition to the improved isolation,
we selected short enough acquisitions windows (which indeed are the most appropriate to the high rates of particles in the DESY beam).
Ddedicated studies in the laboratory are currently pngoingin order to fully understand this issue.

The list of channels mentioned above is discarded from the data taking from now on.

\subsection{Threshold determination}
\label{sec:comm_trigger}

In order to select the optimal trigger threshold values for the detector operation
we perform dedicated scans of trigger threshold values
with all channels enabled (excepted the marked as noisy). 
The threshold values are given in internal DAC units which are translated to
meaningful physical quantities in Section \ref{sec:comm_trigger_sn}. 
The threshold scan curves made of the total number of hits normalized to 1 vs the threshold
for each channel are modeled by a complementary error function called ThS curves from now on:

\begin{equation}
ThS(DAC)=p_{0} \times erfc(\frac{DAC-p_{1}}{p_{2}}) = 
\newline
\frac{2p_{0}}{\sqrt(\pi)} \int_{\frac{DAC-p_{1}}{p_{2}}}^{\infty} e^{-t^{2}} dt,
\label{eq_S-curve}
\end{equation}
where $p_{0}$ is 1/2 of the normalization, $p_{1}$ is the value in which the noise levels are 
the 50\% of its maximum and $p_{2}$ give us the width of the ThS curve.

In the absence of external signals (cosmic rays, injected signals, etc) 
these noise ThS curves show the convolution of the envelope of the 
electronic noise at the output of the fast shaper (the trigger decision branch on the SKIROC).
Indeed, the size of this envelope is related to the slow clock frequency.

To reduce to the minimum the presence of cosmic rays signals, 
we perform runs with short open acquisition windows
as described in the previous section.
In Figure \ref{scurve_channels} 
two result of two threshold scans and the fit by
a ThS curves for two different channels from the second layer of the setup are shown.

\begin{figure}[!t]
  \centering
  \begin{tabular}{ll}
  \includegraphics[width=2.8in]{../figs/commissioning/scurve_chn18_asic7_layer2.eps} & \includegraphics[width=2.8in]{../figs/commissioning/scurve_chn2_asic0_layer2.eps}
  \end{tabular}
\caption{Two threshold scan curves and their associated ThS curves.}
\label{scurve_channels}
\end{figure}

The function from equation \ref{eq_S-curve} was fitted to all
channels data and all the values of $p_{1}$ and $p_{2}$ were saved.
The final threshold value of every ASIC, in DAC units, was chosen 
by taking the

\begin{equation}
maximum(DAC_{optimal}^{ASIC-j} = <p_{1}^{ASIC-j}> + 5 \times <p_{2}^{ASIC-j}>,230)
\end{equation}

if at least the the 30\% of the 64 channels ThS curves in the ASIC could be fitted.
If only less  than the 30\% of curves of the 64 channels were
succesfully fit, a global DAC value of 250 was set.
The $<>$ denotes the average over all the channels on the ASIC.
The choice of these two values was 
sustained by previous experiences, see for example \cite{Amjad:2014tha}.

The optimal trigger threshold values for all ASICs are shown in Figure \ref{trigger_thresholds}.

\subsection{S/N ratio in the trigger line}
\label{sec:comm_trigger_sn}

Similar kind of measurements can be done but using external signals. 
This will allow to calculate the signal over noise (S/N) ratio to trigger 1 MIP signals.
To calculate the S/N of the trigger we need to compare the 1 MIP ThS curve and 2 MIP ThS curve. The S/N 
will be defined as the ratio between the distance of both ThS curves at its 50\% and the width of the 
ThS curve.

In Figure \ref{scurves_injection} we see the 1 MIP and 2 MIP ThS curves obtained for several channel
in a SKIROC testboard in which a single SKIROC2 in BGA package is placed and the 1 MIP and 2 MIPs 
 signals are directly injected in the preamplifier 
(via a 3 pF capacitor located in the injection line as shown in Figure \ref{SKIROC2}). 
From this plot we can extract a S/N ratio of $\sim$12.8. 
We do not expect large differences with the results that we would obtain with 
a full equipped SLABs 
although this board is thought for commissioning and test of the SKIROC ASICS in an
"ideal" environment in contrast with the FEV
ASUS that are optimized to meet the detector requirement and to hold several ASICs at the same time.
%For example: only power pulsing of the analog power is done in this testboard,
%there is a set of external capacitors on the board connected to some ASIC pins and
%usually only one channel is tested. This board allows us to commission and debug the SKIROC itself
%but not the full system.

\begin{figure}[!t]
    \centering
  \begin{tabular}{l}
	\includegraphics[width=4in]{../figs/commissioning/scurve_pp_fastshaper_ch.eps} \\
	\end{tabular}
\caption{ThS curves with charge injection (1 MIP in blue and 2 MIPs in red) for two different channels in a SKIROC2 testboard. From this plot, we extract a $S/N = 12.8$ in the trigger line.}
\label{scurves_injection}
\end{figure}

We have obtained similar results using real signals, in this case cosmic rays signals by using very long
acquisition windows of 150 ms at 5 Hz. This is shown 
in Figure \ref{scurves_cosmics}  where 
we show the result of the fit to the noise ThS curves for all channels individually (in red) in one of the ASICs of 
the second layer together with the results of the ThS curve obtained with cosmic rays integrated for all channels
(black points and blue line).
%For the cosmic ThS curve we took 30 min of data per point using very long spills (150 ms) in 
%frequencies of 5Hz and we add up the triggers from all channels together. In addition, using the 
%knowledge of the MIP value (see next section), we performed a simple cut in the ADC
%to keep only signals corresponding to a MIP 
%(pedestal subtracted) $\pm30$ ADC counts. With this cut we try reduce the contribution of muons 
%traversing the wafers with very large angles but we don't have a way to select perpendicular tracks 
%since we were using simple setups with only single slabs. 
We expect
a broader distribution of the cosmic ThS curve since muons can traverse
the detector at different incidence angles.  
In addition to this, we should remember that the noise ThS curve do correspond to the real noise 
distribution but only to the envelope of the noise in the fast shaper, therefore the
distance between the two ThS curves is smaller than the real distance between noise and signal.
Therefore, if we calculate the S/N 
using this plot, the value would be unrealistic but at least provides 
the comparison between the ThS curve for 1 MIP real and injected signals.

\begin{figure}[!t]
    \centering
  \begin{tabular}{l}
	\includegraphics[width=4in]{../figs/commissioning/cosmic_scurves_asic3_layer2.eps} 
	\end{tabular}
\caption{ThS curves for noise (channel by channel, only the result of the fit) and cosmic rays (all channels together) for one ASIC in layer 2.}
\label{scurves_cosmics}
\end{figure}

Both ways of estimating the S/N ratio of the trigger have their own limitations
and dedicted studies in beam test are needed in order to
precisely determine it.
In the meanwhile, combining the information contained in the Figs \ref{scurves_injection} and \ref{scurves_cosmics},
we can estimate the value in energy at wich we have set our trigger threshold.
This is shown in Figure \ref{trigger_thresholds}.% Two ASICS have a
%trigger threshold value of 0.67 MIP. The 11\% of the ASICs have a trigger threshold between 0.5 and 0.55 MIP and the rest, 87\%, have trigger thresholds slightly lower 0.5 MIP.

\begin{figure}[!t]
  \centering
  \includegraphics[width=4in]{../figs/commissioning/threshold_chip.eps}
  \caption{Summary of the trigger threshold settings in internal DAC units and in MIP units.}
\label{trigger_thresholds}
\end{figure}

\subsection{Prospects}
\label{sec:comm_prospects}

All the commissioning procedure described above relies in very conservative decisions
due to the presence of unknown noise sources during most of the commissioning phase. These sources are
now well know and isolated and therefore a new ``noise commissioning procedure'' has been studied.
It will consist in an iterative algorithm that first
will identify and mask the channels giving underflowed signals and afterwards run a set of acquisitions
in which the number of triggers per channel will be compared with the number of expected triggers
assuming only cosmic rays as signal. This will allow us to have an
unambiguous definition of the noise levels channel per channel instead of 
defining such levels relatively to the total number
of recorded triggers per ASIC. Finally, once that the noisy channels
are identified, the optimization of the threshold levels
will be performed and a last run for identification of noisy channels will be taken using this optimal
threshold.
%In this run, we can decide to mask the noisy channels or
%to increase its own trigger threshold level if we are using Skiroc2a.

Using this new procedure we manage to reduce the number of masked channels by a factor 2 without any loss of performance,
at least in the laboratory and using 3 of the 7 SLABs. This new procedure will be tested in the next beam test.
Also, in order to
optimize the commissioning of the detector,
we propose a new set of measurements in the next beam test such as
a scan of optimal delay of the hold values of the trigger using MIP like particles
and a threshold scan for the determination of the S/N in the trigger line. This later
can be done by the comparison of ThS curves taken with 1 MIP and $\sqrt{2}$ MIP signals (tilting the detector by 45 degrees).

@@

\begin{figure}[!t]
\centering
\begin{tabular}{l}
\includegraphics[width=4.0in]{../figs/proto.png} 
\end{tabular}
\caption{Prototype with 7 layers inside the aluminum stack.}
\label{proto}
\end{figure}

A photograph showing the SiW-ECAL technollogical prototype setup can be seen in Figure \ref{proto}.
Current prototype consists on 7 layers of SLABs housed by a PVC and aluminum structure that can hold up to 10.
For the beam test described in Section \ref{sec:beamtest} all the layers were separated by equal distances of 15 mm
except the last one which was at 60 mm of its nearest. In the following sections, we will refer to layers number 1 to 7, where
the 1 is the closest to the beam pipe and 7 is the farthest.
The detector was exposed to a positron beam in the DESY test beam area (line 24).
By means of an external pulse generator we defined the length of the acquisition window to be
3.7 ms at a frequency of 5 Hz.
The detector was running in power pulsing mode without any extra active cooling system.

@@

First reports on this beam test can be find in
Refs. \cite{Irles:2018uum,Irles:2018hcd}. In this paper we discuss in more detail
the results of the pedestal, noise and MIP calibration in Section \ref{sec:calib}.
We show also results on the pedestal and noise stability when running inside
a magnetic field in Section \ref{sec:magnet}. Finally, a first peek to the 
response and stability of the detector in electromagnetic showers events is discussed in 
Section \ref{sec:showers}.

@@

\begin{figure}[!t]
  \centering
  \begin{tabular}{l}
    \includegraphics[width=4in]{../figs/mip_pedestal_example.eps}
  \end{tabular}
  \caption{Pedestal (blue dashed line) and signal (black continuous line) distribution for one channel in the third layer.}
\label{signal_pedestal}
\end{figure}

>>>>

\usepackage[utf8]{inputenc}
\usepackage{textcomp}
\usepackage{lineno}
\usepackage{color}
\newcommand{\todo}[1]{\textcolor{red}{{#1}}}
\usepackage[outdir=./]{epstopdf}

@@

\begin{frontmatter}
  \title{Beam test performance of the highly granular SiW-ECAL technological prototype for the ILC.}

@@

\end{frontmatter}

\linenumbers

@@

The entity
of sensors, thin PCB (printed circuit boards) and ASICs (application-specific integrated circuits)
is called Active Signal Units or ASU.
An individual ASU has a lateral dimension of 18x18 cm$^{2}$.
The ASUs are currently equipped
further with 16 SKIROC2\cite{}  (Silicon pin Kalorimeter Integrated ReadOut Chip)
ASICs for the read out and features 1024 square pads (64 per ASIC) of 5x5 mm.
The channels and ASICs are distributed along the ASU as shown in Figure \ref{ASU}.
Each ASU is equipped with 4 silicon wafers. The sensors consist on floating zone silicon wafers
320$\mu$m thick with high resistivity (bigger than 5000 $\Omega\cdot$cm).
The size of the wafer is $9\times9$ cm$^{2}$ and it is subdivided in an array of 256 PIN diodes of $5\times5$ mm$^{2}$.
A MIP (minimum ionizing particle) traversing the PIN parallel to its normal will create $\sim$ 80 $h^{+}e^{-}$ pairs per $\mu$m which corresponds to 4.1 fC
for particles incident perpendicularly to its surface.
The high voltage is delivered to the wafers using a high voltage kapton sheet that covers the full extension of the wafers.
The readout layers of the SiW-ECAL consist of a chain of Active Sensor Units (ASUs) and an adapter board
to a data acquisition system (DAQ) at the beginning of the layer.
The readout layers are embedded on a "U" shape carbon structure to protect the wafers.
The full system is then covered by two aluminum plates
to provide electromagnetic shielding and mechanical stability.
This ensemble is denominated SLAB
(``short'' for 1 ASU ensembles or ``long'' for several ASUs enchained) and it can be seen in
Figure \ref{ASU2}.
With current SLABs, a potential density of
4000 channels/dm$^{3}$ is achievable. This number should be compared with
the density achieved for previous beam tests: 1500 channels/dm$^{3}$ \cite{Amjad:2014tha}.

\begin{figure}[!t]
  \centering
    \includegraphics[width=0.4\textwidth]{../figs/test.jpg} 
  \caption{\todo{Temporary picture: La Esperanza del Condenado, J. Miró}: Open single SLAB with FEV11 ASU, 16 SKIROC 2, interface card and DIF visibles. }
\label{ASU2}
\end{figure}

\begin{figure}[!t]
  \centering
  \begin{tabular}{l}
    \includegraphics[width=0.4\textwidth]{../figs/ASU_geometry1.eps}  \\
    \includegraphics[width=0.4\textwidth]{../figs/ASU_geometry2.eps}  \\
  \end{tabular}
  \caption{Repartition of the ASIC (up) and channels (down) in one ASU. In this perspective, the Si-Sensors are in glued in the back.
    The channels are separated (in x and y) by 5.5 mm.
    The empty cross in the middle of the ASU corresponds to the 1 mm separation between the sensors.
    The areas covered by the different ASICs and channels
    are labeled with numbers following design and DAQ criteria: from 0-16 in the case of the ASICs and from 0-63 in the case of the channels.
  }
\label{ASU}
\end{figure}

The SiW-ECAL detector designed for the ILD requires of the order
of $10^{5}$ highly integrated detection like the ones described in this text.
For the production of the small sample of SLABs studied in this document,
a scalable working procedure has been established among several groups \cite{Boudry:2318814}
profiting from the funding of projects like AIDA2020 or the HIGHTEC emblematic project
of the P2IO.

The subsequent chain of the data acquisition (DAQ) system \cite{Gastaldi:2014vaa} is inspired by the ILC.
It consists on three layers of modules.
The first layer, or most downstream layer, consists on a set of detector interfaces called DIFs which are placed at the beginning of each SLAB holding up to 15 ASUs.
The second layer is made of concentrator cards module called the Gigabit Concentrator Cards (GDCCs)
which are used to control up to 7 DIFs collecting all data from them and distributing among them the system clock and fast commands.
The last layer consist on a clock and control card (CCC) which
provides a clock, control fan-out of up to 8 GDCCs and accepts and distributes external signals (i.e. signals
generated external pulse generator to simulate the ILC spill conditions).
The whole system is controlled by the Calicoes and the Pyrame DAQ software version 3~\cite{Rubio-Roy:2017ere,Magniette:2018wdz}.

\begin{figure}[!t]
\centering
\begin{tabular}{l}
\includegraphics[width=0.4\textwidth]{../figs/proto.png} 
\end{tabular}
\caption{Prototype with 7 layers inside the aluminum stack.}
\label{proto}
\end{figure}

A photograph showing the SiW-ECAL technollogical prototype setup can be seen in Figure \ref{proto}.
Current prototype consists on 7 layers of SLABs housed by a PVC and aluminum structure that can hold up to 10.
For the beam test described in Section \ref{sec:beamtest} all the layers were separated by equal distances of 15 mm
except the last one which was at 60 mm of its nearest. In the following sections, we will refer to layers number 1 to 7, where
the 1 is the closest to the beam pipe and 7 is the farthest.
The detector was exposed to a positron beam in the DESY test beam area (line 24).
During the full data taking period at the DESY beam test facility,
the detector was running in power pulsing mode without any extra active cooling system.
and means of an external pulse generator we defined the length of the acquisition window to be
3.7 ms at a frequency of 5 Hz.

@@

@@

First reports on this beam test can be find in
Refs. \cite{Irles:2018uum,Irles:2018hcd}. In this paper we discuss in detail
the results of the pedestal, noise and MIP calibration in Section \ref{sec:calib}.
We also show the results on the pedestal and noise stability when running inside
a magnetic field in Section \ref{sec:magnet}. The study of the perfomance of the detector
for electromagnetic shower events will be covered in future publciations.

\subsection{Commissioning}
\label{sec:commissioning}

Earlier experiences with the SKIROC2 ASIC are reported in Refs. \cite{Amjad:2014tha,Suehara:2018mqk}). 
Internal SKIROC2 parameters found in these references are adopted in the following
except if the opposite is stated.
For example, the gain value of 1.2pF for the preamplifier is used. 
With this gain, the SKIROC2 ensures a linearity better than 90\% 
for 0.5-200 MIPs, which is enough for 
electromagnetic showers created by few GeV 
electrons or positrons.
This beam test was prepared by a careful and comprehensive commissioning comprising
the debug of the
short SLABSs with special emphasis in the control of the noise and the study of the
prototype performance in cosmic rays tests. Dedicated runs at the laboratory were taken
in order to identify and mask the noisy channels. The outcome of this study
is that X\% of the channels were masked before putting the detector in front of the beam.
Two different kinds of noisy channels were found: the ones that are noisy on all
SLABs and those which are only noisy in one or few of them. Preliminary inspections
of the PCB hints that the first group is associated to routing issues on the PCB and
hereafter they are labelled like this. The summary of this study is shown in Figure \ref{noisycells}.
These channels are removed from the data taking in all the studies shown from now on.

\begin{figure}[!t]
  \centering
  \begin{tabular}{l}
  \includegraphics[width=0.4\textwidth]{../figs/commissioning/masked_layer.eps} \\
  \includegraphics[width=0.4\textwidth]{../figs/commissioning/masked_chip.eps}
  \end{tabular}
\caption{Ratio of channels that are marked as noisy in all slabs. 
Top: inventory of the different type of noisy channels per slab. 
Bottom: break down of the total number of noisy channels per ASIC. 
The ASICs 4-7 (wafer issue) and 10 from layer 1 and the ASIC 4 from layer 7 are not included
in the second plot since they are fully masked.}
\label{noisycells}
\end{figure}

\subsubsection{Trigger threshold optimization}
\label{sec:comm_trigger}

\todo{Autotrigger stuf.
In order to select the optimal trigger threshold values for the detector operation
we perform dedicated scans of trigger threshold values
with all channels enabled (excepted the marked as noisy). }
The optimal trigger threshold values for all ASICs are shown in Figure \ref{trigger_thresholds}.

\begin{figure}[!t]
  \centering
  \includegraphics[width=0.4\textwidth]{../figs/commissioning/threshold_chip.eps}
  \caption{Summary of the trigger threshold settings in internal DAC units and in MIP units.}
\label{trigger_thresholds}
\end{figure}

\subsubsection{S/N ratio in the trigger line}
\label{sec:comm_trigger_sn}

In Figure \ref{scurves_injection} we see threshold scans curves
 curves obtained for several channel
in a SKIROC testboard in which a single SKIROC2 in BGA package is placed and the 1 MIP and 2 MIPs 
 signals are directly injected in the preamplifier 
(via a 3 pF capacitor located in the injection line as shown in Figure \ref{SKIROC2}). 
From this plot we can extract a S/N ratio of $\sim$12.8. 
Unfortunately, this results has obtained using
a dedicated board thought for commissioning and test of the SKIROC ASICS in an
"ideal" environment in contrast with the fully equipped
ASUS that are optimized to meet the detector requirement and to hold several ASICs at the same time.
We do not expect large differences with the results that we would obtain with 
a full equipped SLABs but dedicated runs in beam test will be needed.

\begin{figure}[!t]
    \centering
  \begin{tabular}{l}
	\includegraphics[width=0.4\textwidth]{../figs/commissioning/scurve_pp_fastshaper_ch.eps} \\
	\end{tabular}
\caption{Threshold curves with charge injection (1 MIP in blue and 2 MIPs in red) for two different channels in a SKIROC2 testboard. From this plot, we extract a $S/N = 12.8$ in the trigger line. \todo{We need to use charge in the axis}}
\label{scurves_injection}
\end{figure}

@@

\begin{figure}[!t]
  \centering
  \begin{tabular}{l}
    \includegraphics[width=0.4\textwidth]{../figs/mip_pedestal_example.eps}
  \end{tabular}
  \caption{Pedestal (blue dashed line) and signal (black continuous line) distribution for one channel in the third layer.}
\label{signal_pedestal}
\end{figure}

The pedestal is calculated as the mean position of
the ADC distribution of channels without trigger. The noise is
associated to the width of such distribution.
The pedestal correction is done layer-, chip-, channel- and SCA-wisely due to the large spread of values between pedestals, as observed in 
Figure \ref{pedestal_layer} (upper plot) and Figure \ref{pedestal_all} (also upper plot).
For the noise, the dispersion is much smaller ($\sim 5 \%$). This is shown in the lower plots of Figures \ref{pedestal_layer} and \ref{pedestal_all}.
From now on, the pedestal correction is applied to all the results presented.
